---
title: "Explore One Variable Problem Set"
author: "Justin Le"
date: "October 26, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(gridExtra)
data("diamonds")
summary(diamonds)
```

***

### Looking into the peak

```{r}
p1 <- ggplot(aes(x = price), data = diamonds) +
  geom_histogram(binwidth = 250, color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(limits = c(0, 20000), breaks = seq(0, 20000, 2000))

p2 <- ggplot(aes(x = price), data = diamonds) +
  geom_histogram(binwidth = 25, color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(limits = c(0, 2000), breaks = seq(0, 2000, 250))

grid.arrange(p1, p2, ncol = 1)
```

```{r}
summary(diamonds$price)

subset(diamonds, price < 500)
subset(diamonds, price < 250)
nrow(subset(diamonds, price >= 15000))
```
***

### Histogram of diamond prices by cut

```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(color = I('black'), fill = I('#099DD9')) +
  facet_wrap(~cut)
```

***

### Price by cut questions:
Which cut has the highest priced diamond? Premium
Which cut has the lowest priced diamond? Premium, Ideal
Which cut has the lowest median price? Ideal

```{r}
by(diamonds$price, diamonds$cut, summary)

tapply(diamonds$price, diamonds$cut, max)
tapply(diamonds$price, diamonds$cut, min)
tapply(diamonds$price, diamonds$cut, median)
```

***

### Don't want y-axis to be fixed

```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(color = I('black'), fill = I('#099DD9')) +
  facet_wrap(~cut, scales = "free_y")
```

***

### Histograms of price per carat and facet it by cut

```{r}
ggplot(aes(x = price/carat), data = diamonds) +
  geom_histogram(binwidth = 0.05, color = I('black'), fill = I('#FFA500')) +
  scale_x_log10() +
  facet_wrap(~cut, scales = "free_y")
```

***

### Investigate the price of diamonds using box plots,
numerical summaries, and one of the following categorical
variables: cut, clarity, or color.

```{r}
ggplot(aes(x = color, y = price), data = diamonds) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 8000))
```

```{r}
by(diamonds$price, diamonds$color, summary)
```

What is the IQR for the best color? 3302.5
IQR for the worst?

```{r}
IQR(subset(diamonds, color == 'D')$price)
IQR(subset(diamonds, color == 'J')$price)
```

***
### Investigate the price per carat of diamonds across
the different colors of diamonds using boxplots.

```{r}
ggplot(aes(x = color, y = price/carat, fill = color), data = diamonds) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 6500))
```

```{r}
by(diamonds$price/diamonds$carat, diamonds$color, summary)
```

Look what happens when we look at only the heavier (greater than median carat) diamonds with the best cut and best clarity. This is more what we would expect, as color gets worse price per carat decreases. The reason for this maybe because people are will to pay more for bigger diamond even if they have a bad cut, bad clarity and bad color. 
Agree?? Disagree?? 
Wow, over $15,000 per carat for a diamond with the best cut, clarity and color!

```{r}
uniform <- subset(diamonds, cut == "Ideal" & clarity == "IF" & carat > 0.7)
qplot(x = color, y = price/carat, data = uniform, geom = "boxplot", fill = color)
```

***

### Investigate the weight (carat) of the diamonds using a frequency polygon. Use
different binwidths to see how the frequency changes. 

```{r}
ggplot(aes(x = carat), data = diamonds) +
  geom_freqpoly(binwidth = 0.01) +
  coord_cartesian(xlim = c(0, 2.5))
```

### What carat size has a count greater than 2000?

```{r}
table(diamonds$carat)
```

You can save images by using the ggsave() command.
ggsave() will save the last plot created.
For example...
                 qplot(x = price, data = diamonds)
                 ggsave('priceHistogram.png')

ggsave currently recognises the extensions eps/ps, tex (pictex),
pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

***
