---
title: "Explore One Variable"
author: "Justin Le"
date: "`r format(Sys.Date())`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook_User_Data}
getwd()
setwd('D:/Udacity Data Analysis with R/Lesson 3')
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram_of_Users\'_Birthdays}
library(ggplot2)

names(pf)
#qplot(x = dob_day, data = pf) +
#  scale_x_discrete(breaks=1:31)
#Error: StatBin requires a continuous x variable the x variable is discrete. Perhaps you want stat="count"?

ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***

### Faceting
Notes:

```{r Faceting}
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)
```

***

#### What code would you enter to create a histogram of friend counts?

```{r Friend_Count}
#qplot(x = friend_count, data = pf)

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```

***

### Limiting the Axes
Notes:

```{r Limiting_the_Axes}
#qplot(x = friend_count, data = pf, binwidth = 25) +
#  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

***

### Faceting Friend Count

```{r Faceting_Friend_Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting_NA_Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics_\'by\'_Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response: Because outliers would pull the mean far to the right, skewing the data. Median is a more robust datapoint.

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, binwidth = 30, color = I('black'), fill = I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure_Histogram_by_Year}
qplot(x = tenure/365, data = pf, binwidth = .25, color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7))
```

***

### Labeling Plots
Notes:

```{r Labeling_Plots}
qplot(x = tenure/365, data = pf, binwidth = .25,
      xlab = 'Number of Years Using Facebook',
      ylab = 'Number of Users in Sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7))
```

***

### User Ages
Notes:

```{r User_Ages}
ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 1, fill = '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5))
```

***

### Transforming Data
Notes:
***

### Create multiple plots on one page
Notes: install.packages("gridExtra")
  # define individual plots
    p1 = ggplot(...)
    p2 = ggplot(...)
    p3 = ggplot(...)
    p4 = ggplot(...)
  # arrange plots in grid
    grid.arrange(p1, p2, p3, p4, ncol=2)
    
```{r}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)
summary(log10(pf$friend_count + 1))
summary(sqrt(pf$friend_count))
```

### Transforming Data Solution

```{r}
library(gridExtra)

p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count + 1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)

grid.arrange(p1, p2, p3, ncol = 1)
```

***

### Add a Scaling Layer
Notes:

```{r Add_a_Scaling_Layer}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```

***

### Frequency Polygons

```{r Frequency_Polygons}
qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count',
      ylab = 'Proportion of Users with that Friend Count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))

ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***

### Likes on the Web
Notes:

What's the www_like count for males? --> 1430175

Which gender has more www_likes? --> Females

```{r}
by(pf$www_likes, pf$gender, sum)
```


```{r Likes_on_the_Web}
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()
```

***

### Box Plots
Notes:

```{r Box_Plots}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.
Notes:scale_x/y_continuous clips (removes unseen data points) outside of given range
coord_cartesian (preferred) does not remove data points from calculations

```{r}
ggplot(aes(x = gender, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 1000))

ggplot(aes(x = gender, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box_Plots_Quartiles_and_Friendships}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response: women 

```{r Friend_Requests_by_Gender}
ggplot(aes(x = gender, y = friendships_initiated),
       data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 150))

by(pf$friendships_initiated, pf$gender, summary)
```

***

### Getting Logical
Notes: Sometimes when there's a variable with a lot of zeros you may want to transform the data into something more usable.
Here we take mobile logins. Using the raw data isn't that useful. So we can ask the question of how many have ever used mobile login. We can
transform the data into binary.

```{r Getting_Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection: